---
title: "stats101a_hw7"
author: "Jaden Chou - 206654728"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(tidyverse)
library(ggplot2)
```


## (1)

```{r}
df <- data.frame(read.csv('./AdRevenue.csv'))

str(df)
```


Part A

### (a)

```{r}
model <- lm(AdRevenue ~ Circulation, df)

summary(model)
```
```{r}
plot(model)
```

```{r}
log_model <- lm(AdRevenue ~ I(log(Circulation)), df)

summary(log_model)
```
```{r}
plot(log_model)
```
log(Circulation) seems worse. Especially residual vs. fitted plot where there's an identifiable trend now. R-squared value is lower as well.

```{r}
log_both_model <- lm(I(log(AdRevenue)) ~ I(log(Circulation)), df)

summary(log_both_model)
```
```{r}
plot(log_both_model)
```

This model is much better. Residuals vs. fitted is much better and there's a lack of a trend. QQ-plot seems much better as well where most values are in a straight line. Scale-location seems better. No bad high leverage either.

### (b)

```{r}
log_pred <- predict(log_both_model, newdata = data.frame(Circulation = c(0.5, 20)), interval = 'prediction', level = 0.95)

exp(log_pred)
```

74.30864 with (51.82406, 106.5485)
522.56626 with (59.89585 758.7626)

### (c)

predictions become less intuitive, intervals are multiplicative and asymmetric. Constant variance of errors seem a bit iffy with the scale-location plot.

Part B

```{r}
model1 <- lm(AdRevenue ~ Circulation, data = df)
model2 <- lm(AdRevenue ~ Circulation  + I(Circulation^2), data = df)
model3 <- lm(AdRevenue ~ Circulation + I(Circulation^2) + I(Circulation^3), data = df)
```

```{r}
summary(model1)

plot(model1)
```
```{r}
summary(model2)

plot(model2)
```
This one is no good. The upward-trend seen in the Scale-Location plot indicates that it is not valid in terms of the constant variance of errors. The QQ-plot is worse as well.

```{r}
summary(model3)

plot(model3)
```
Scale-location is off again. QQ-plot is bad as well. Residuals plot ain't good either.

So the best choice of model is still the simple linear model without any polynomial transformations.

### (b)

```{r}
predict(model1, newdata = data.frame(Circulation = c(0.5, 20)), interval = 'prediction', level = 0.95)
```


111.2362 with (26.2383, 196.2342)
556.8774 with (466.1694, 647.5854)

### (c)

There are outliers in the simple regression model as well as a residual plot that is not the best (all the values are congregated very closely at low fitted values)

Part C

### (a)

Since the model in Part B is just the normal simple regression model, and since we compared that directly with our chosen model in Part A, then the model in Part A is better (double log transform), with reasons explaiend above. Examined the validiity plots for the most part.

### (b)

Part A:
74.30864 with (51.82406, 106.5485)
522.56626 with (59.89585 758.7626)

Part B:
111.2362 with (26.2383, 196.2342)
556.8774 with (466.1694, 647.5854)

I would recommend the bottom one (Part B) just because it's more intuitive and easy to interpret. Since the intervals in Part A were logged and then exponentiated, it will be harder to interpret.

## (2)

```{r}
df <- data.frame(read.csv('./marioKart.csv'))

str(df)
```

### (a)

```{r}
model <- lm(totalPr ~ startPr + shipPr + duration + nBids + sellerRate + wheels, data = df)

summary(model)
```

Holding the rest of the predictors constant, a 1-unit increase in the number of bids is associated with an average change of +0.2375 dollars in the total price.

### (b)

Holding all other variables constant, one more wheel is associated with an average change of +7.938 dollars in the total price.

### (c)

```{r}
model2 <- lm(totalPr ~ startPr + shipPr + duration + nBids + sellerRate + wheels + cond, data = df)

summary(model2)
```

Holding all variables constant, a used game is associated with an average change of -4.45 dollars in total price compared to a new game. Yes, it does matter with all things being equal -- it is worth 4.45 dollars less on average.

## (3)

```{r}
df <- data.frame(read.csv('./practiceproblem.csv'))

str(df)

y <- df$finish_time
x <- df$barrier_time
resids <- df$resids
```
### (a)

```{r}
# SXX = variation of x
SXX <- sum((x - mean(x))^2)
SXX
```

### (b)

```{r}
# SYY = variation of y
SYY <- sum((y - mean(y))^2)
SYY
```

### (c)

```{r}
# estimate = s^2 = SYY / (n - 1)

var <- SYY / (length(df) - 1)

var
```

### (d)

```{r}
## SSY = SSreg + RSS; SSreg = SSY - RSS

SSY <- SYY
RSS <- sum(resids^2)

SSreg <- SSY - RSS

SSreg

```

### (e)

```{r}
## R-squared = SSreg / SSY

R_squared <- SSreg / SSY

R_squared
```

### (checking)

```{r}
model <- lm(y ~ x)

summary(model)
anova(model)
```



