---
title: "stats101a_hw9"
author: "Jaden Chou - 206654728"
date: "`r Sys.Date()`"
output: pdf_document
---


```{r}
library(tidyverse)
library(ggplot2)
library(car)
library(leaps)
```

## (1)

```{r}
df <- data.frame(read.csv('./pgatour2006-3.csv'))

str(df)
```
```{r}
df <- df |>
  select(-TigerWoods, -Name)
```

```{r}
str(df)
```

### (a)

```{r}
best_fit <- regsubsets(
  PrizeMoney ~ .,
  data = df,
  nvmax = ncol(df) - 1
)

best_summary <- summary(best_fit)

best_model_index <- which.max(best_summary$adjr2)
best_summary$which[best_model_index, ]

```
```{r}
final_model <- lm(
  PrizeMoney ~ DrivingAccuracy +
                GIR +
                BirdieConversion +
                SandSaves +
                Scrambling,
  data = df
)

summary(final_model)
```

$\text{PrizeMoney} = \beta_0 + \beta_1(\text{DrivingAccuracy}) + \beta_2(GIR) + \beta_3(\text{BirdieConversion}) + \beta_4(\text{SandSaves}) + \beta_5(\text{Scrambling}) + \epsilon$

### (b)

```{r}
best_fit <- regsubsets(
  PrizeMoney ~ .,
  data = df,
  nvmax = ncol(df) - 1
)

best_summary <- summary(best_fit)

best_bic_index <- which.min(best_summary$bic)
best_summary$which[best_bic_index, ]
```
$\text{PrizeMoney} = \beta_0 + \beta_1(\text{DrivingAccuracy}) + \beta_2(\text{GIR}) + \beta_3(\text{BirdieConversion}) + \beta_4(\text{Scrambling}) + \varepsilon$

This one does not include the `Scrambling` feature.


### (c)

```{r}
forward_fit <- regsubsets(
  PrizeMoney ~ .,
  data = df,
  method = "forward",
  nvmax = ncol(df) - 1
)

forward_summary <- summary(forward_fit)
```

```{r}
best_adjr2 <- which.max(forward_summary$adjr2)

forward_summary$which[best_adjr2, ]
```
Best model using adjusted R-squared:

$\text{PrizeMoney} = \beta_0 + \beta_1(\text{DrivingAccuracy}) + \beta_2(\text{GIR}) + \beta_3(\text{BirdieConversion}) + \beta_4(\text{SandSaves}) + \beta_5(\text{Scrambling}) + \varepsilon$

```{r}
best_bic <- which.min(forward_summary$bic)

forward_summary$which[best_bic, ]
```
Best model using BIC:

$\text{PrizeMoney} = \beta_0 + \beta_1(\text{DrivingAccuracy}) + \beta_2(\text{GIR}) + \beta_3(\text{BirdieConversion}) + \beta_4(\text{Scrambling}) + \varepsilon$

### (d)

```{r}
backward_fit <- regsubsets(
  PrizeMoney ~ .,
  data = df,
  method = "backward",
  nvmax = ncol(df) - 1
)

backward_summary <- summary(backward_fit)

best_adjr2 <- which.max(backward_summary$adjr2)

backward_summary$which[best_adjr2, ]
```
Best model using adjusted R-squared:

$\text{PrizeMoney} = \beta_0 + \beta_1(\text{DrivingAccuracy}) + \beta_2(\text{GIR}) + \beta_3(\text{BirdieConversion}) + \beta_4(\text{SandSaves}) + \beta_5(\text{Scrambling}) + \varepsilon$

```{r}
best_bic <- which.min(backward_summary$bic)

backward_summary$which[best_bic, ]
```
Best model using BIC:

$\text{PrizeMoney} = \beta_0 + \beta_1(\text{DrivingAccuracy}) + \beta_2(\text{GIR}) + \beta_3(\text{BirdieConversion}) + \beta_4(\text{Scrambling}) + \varepsilon$

### (e)

variables that repeatedly appear:
 - DrivingAccuracy
 - GIR
 - BirdieConversion
 - Scrambling

some:
 - SandSaves
 

## (2)

### (a)

```{r}
full_model <- lm(PrizeMoney ~ ., data = df)
summary(full_model)
```

```{r}
par(mfrow = c(2, 2))

plot(full_model)
```

There's a trend in residuals v. fitted that shows breaks the linear trend.
constant variance is broken (fanning in residuals v. fitted and increasing trend in scale-location)
QQ plot is not the best, errors are not standard.

### (b)

```{r}
invResPlot(full_model)
```

log transform is quite close to the optimal lambda.

```{r}
summary(powerTransform(full_model))
```

the recommended rounded power is 0. With the LR test, lambda = 0 p value > 0.05, so we do not reject. The LR test for lambda 1 has pvalue of < 2.22e-16 < 0.05, so we reject the null hypothesis , concluding we do need a transformation.

log transform is good.

### (c)

```{r}
log_model <- lm(log(PrizeMoney) ~ ., data = df)

par(mfrow = c(2, 2))

plot(log_model)
```

Validity is improved. There is a much lesser indication of a trend in Residuals v. Fitted. And constant variance seems to be good (or at the very least much better) in residuals vs. fitted and scale-location plot. Q-Q residuals looks much more straight line as well.

### (d)

```{r}
pt_pred <- powerTransform(
  cbind(AveDrivingDistance,
        DrivingAccuracy,
        GIR,
        PuttingAverage,
        BirdieConversion,
        SandSaves,
        Scrambling,
        BounceBack,
        PuttsPerRound) ~ 1,
  data = df
)

summary(pt_pred)
```
```{r}
df_trans <- df

df_trans$AveDrivingDistance <- bcPower(df$AveDrivingDistance, pt_pred$roundlam[1])
df_trans$DrivingAccuracy    <- bcPower(df$DrivingAccuracy,    pt_pred$roundlam[2])
df_trans$GIR                <- bcPower(df$GIR,                pt_pred$roundlam[3])
df_trans$PuttingAverage     <- bcPower(df$PuttingAverage,     pt_pred$roundlam[4])
df_trans$BirdieConversion   <- bcPower(df$BirdieConversion,   pt_pred$roundlam[5])
df_trans$SandSaves          <- bcPower(df$SandSaves,          pt_pred$roundlam[6])
df_trans$Scrambling         <- bcPower(df$Scrambling,         pt_pred$roundlam[7])
df_trans$BounceBack         <- bcPower(df$BounceBack,         pt_pred$roundlam[8])
df_trans$PuttsPerRound      <- bcPower(df$PuttsPerRound,      pt_pred$roundlam[9])
```

```{r}
bc_model <- lm(
  PrizeMoney ~ .,
  data = df_trans
)

par(mfrow = c(2, 2))

plot(bc_model)
```

Linearity trend is violated as there exists a trend in the residuals vs fitted plot. Constant variance is violated (fanning in residuals plot and upward trend in scale-location plot). QQ residuals is slightly better but still not amazing.

### (e)

```{r}
log_bc_model <- lm(
  log(PrizeMoney) ~ .,
  data = df_trans
)

par(mfrow = c(2, 2))

plot(log_bc_model)
```

I recommend the simply y-transformed model. The y-transformed model and the y and x's transformed model seem very similar, with both of them being much better than the other two models. But, the y-transformed model is easier to intrepret compared to the x's being transformed as well, which is why I recommend that one -- there's not a large enough improvement in validity for it to be worth the extra complications in interpretation.

### (f)

```{r}
hat_values <- hatvalues(log_model)

n <- nrow(df)
p <- length(coef(log_model))

high_lev <- which(hat_values > 2 * p / n)
high_lev
```

28  40 168 178 are observations worth investigating.

```{r}
cooks <- cooks.distance(log_model)

influential <- which(cooks > 4 / n)
influential
```

  1   9  40  47  60  63  84 101 122 125 128 168 180 185 are observations worth investigating.
  
  