---
title: "stats101a_hw8"
author: "Jaden Chou - 206654728"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(tidyverse)
library(ggplot2)

install.packages('car', repos = "https://cran.r-project.org")
library(car)
```

## (1)

```{r}
df <- data.frame(read.csv('./cars04.csv'))

str(df)
```
```{r}
numeric_vars <- df[, sapply(df, is.numeric)]
numeric_vars <- numeric_vars[, !names(numeric_vars) %in% c("Hybrid")]


model <- lm(SuggestedRetailPrice ~ ., data = numeric_vars)
summary(model)
```

### (a)

`Vehicle.name` is just the name of the vehicle -- it's an identifier similar to ids or keys that provide information to us, the user, but does not provide much information to a model. That's why we exclude it from the model -- it does not tell us anything important about `SuggestedRetailPrice` beyond its features / specs that are already extrapolated with the rest of the predictors.

### (b)

$$\widehat{\text{SuggestedRetailPrice}} = \beta_0 + \beta_1(\text{DealerCost}) + \beta_2(\text{EngineSize}) + \beta_3(\text{Cylinders}) + \beta_4(\text{Horsepower}) + \beta_5(\text{CityMPG}) + \beta_6(\text{HighwayMPG}) + \beta_7(\text{Weight}) + \beta_8(\text{WheelBase}) + \beta_9(\text{Length}) + \beta_{10}(\text{Width})$$
### (c)

```{r}
summary(model)
```

slope coefficient: 228.32952
t-value: 3.171
p value: 0.001730

Since the p-value = 0.001730 < $\alpha$ = 0.05, there is statistically significant evidence to reject the null hypothesis ($H_0: B_{\text{cylinders}} = 0$). That is, there is evidence that the number of cylinders has a linear relationship with suggested retail price, controlling for all other predictors. 

### (d)

For each additional cylinder, the suggested retail price increases by approximately $228.33 on average, all other predictors being the same.

### (e)

```{r}
anova(model)

# get F value: 9.6084e+00
```
```{r}
# t = sqrt(F)

t <- sqrt(9.6084e+00)

t
```

### (f)

```{r}
summary(model)
```

F-statistic = 2.073e+04

The F-statistic tests the overall significance of the model, that is:

$H_0: \beta_{1} = \beta_{2} =  ...  = \beta_{10} = 0$

$H_a: \text{at least one } \beta_i \neq 0$

Since the p-value < 2.2e-16 < 0.05, we reject $H_0$ and conclude that at least one predictor has a statistically significant linear relationship with suggested retail price. So the model as a whole is statistically significant.

## (2)

```{r}
df <- data.frame(read.csv('./pgatour2006-3.csv'))

str(df)
```

```{r}
df <- subset(df, select = -c(Name, TigerWoods))

str(df)
```
```{r}
model <- lm(PrizeMoney ~ ., data = df)
summary(model)
```

### (a)

```{r}
pairs(df)
```

`PuttingAverage` and `PuttsPerRound`.

`DrivingDistance` and `DrivingAccuracy`

`PuttingAverage` and `BirdieConversion`

### (b)

```{r}
vif(model)[vif(model) > 5]
```

`PuttingAverage`, `BirdieConversion`, and `PuttsPerRound` are greater than 5, which relates to the issues with collinearity that I saw in (a). Interestingly, and even after investigating the scatterplots in hindsight, `GIR` has a VIF value > 5 but I don't really think the collinearity is too big with it.

### (c)

An inflated standard error directly inflates the denominator of the t-statistic since

$$t = \frac{\hat{\beta}_j}{SE(\hat{\beta}_j)}$$
So a larger Standard Error results in a smaller magnitude t score which results in a larger p-value. This means the variable may appear statistically insignificant not because it truly has no effect, but because collinearity is masking it.

### (d)

```{r}
par(mfrow = c(2,2))
plot(model)
```

Non-constant variance:
 - residuals v fitted plot show a fan pattern where variance increases with fitted values
 - scale-location plot has strong upward trend

non-normal:
 - QQplot has significant deviation from the straight line at the upper tail.
 
trend/non-linearity:
- residuals vs fitted plot shows a clear pattern/trend where it decreases in a straight line then goes back up .

### (e)

```{r}
log_model <- lm(log(PrizeMoney) ~ ., data = df)

par(mfrow = c(2,2))
plot(log_model)
```

Yes a log transform of the response variable makes the validity much better.

The constant variance is much better supported. The QQplot is much more following hte line. Lack of a trend in residuals v. fitted as well as pretty constant variacnce.

### (f)

```{r}
n <- nrow(df)
p <- 9  
threshold <- 2*(p+1)/n
threshold

# get leverage values
leverage <- hatvalues(model)
which(leverage > threshold)
```

potential bad high leverage points = observations 28 40 168 178

### (g)

```{r}
cooks <- cooks.distance(model)
which(cooks > 1)
```

observation 178 is a high influence point.