---
title: "stats101a_hw10"
author: "Jaden Chou - 206654728"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(tidyverse)
```
```{r}
df <- data.frame(read.csv('./dietstudy.csv'))

str(df)
```

## prepare data

```{r}
df <- df |>
  select(DIET, SEX, WEIGHT_0, WEIGH_12, AVEAD_12)
```

```{r}
df <- df |>
  mutate(weight_change = WEIGH_12 - WEIGHT_0)

df |> slice_sample(n = 2)
```

```{r}
df <- df |> select(-WEIGH_12) 

```

## (Q1)

```{r}
model <- lm(weight_change ~ ., data =df)

summary(model)
```

### (a)

holding all other variables constant, there is no evidence that diet type does affect weight change. The estimated differences between diets are relatively small relative to standard errors, and the p-values are high (> 0.05), so not statistically significant.

```{r}
reduced <- lm(weight_change ~ SEX + WEIGHT_0 + AVEAD_12, data = df)

anova(reduced, model)
```
Because p = 0.9017 which is much larger than 0.05, we fail to reject the null hypothesis $H_0$ that diet is associated with weight change. There is no evidence that diet type contributes to explaining weight change after accounting for sex, baseline weight, and average adherence

### (b)

Yes, the p-value (1.15e-13) is far smaller than 0.05, so we reject $H_0$ and can conclude with very strong evidence that adherence is associated with weight change, controlling for other variables.

The estimated coefficient is -1.374, for each one-unit increase in average adherence, the expected weight change decreases by about 1.374 kg, all other variables held constnat.

### (c)

No. Since the p-value for SEXMale is 0.6544 > 0.05, we fail to reject $H_0$ -- holding all variables constant, there is no statistically significant evidence that males and females differ in average weight change.

### (d)

The WEIGHT_0 parameter of -0.056 indicates that for every 1 kg that a person initially started at, they would lose an estimated 0.056 more kg in weight after the 12 months.

## (2)

```{r}
model_interaction <- lm(weight_change ~ WEIGHT_0 + SEX * AVEAD_12, data = df)

summary(model_interaction)
```
Since the p-value of the interaction term is 0.233 ( > 0.05), we fail to reject $H_0$.

There is no statistically significant evidence that the effect of adherence on weight change differs between males and females.

## (3)

```{r}
diets <- df |> mutate(weight0_centered = WEIGHT_0 - mean(WEIGHT_0, na.rm = TRUE)) |> select(-WEIGHT_0)

str(diets)
```
```{r}
model <- lm(weight_change ~ ., data = diets)

summary(model)
```
The expected (average) weight change for a female on the Atkins diet whose initial weight is equal to the sample mean initial weight and whose average adherence is 0 would be about 3 kg.


