---
title: "takehome_final_exam"
author: "Jaden Chou - 206654728"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(car)
library(MASS)
library(tidyverse)

library(leaps)
```

## (1)

```{r}
df <- data.frame(read.csv('./crabs.csv'))
```

```{r}
df1 <- df |> dplyr::select(-Sex, -size)

str(df1)
```

```{r}
model <- lm(PinchingForce ~ ., data = df1)

par(mfrow = c(2,2))

plot(model)
```

```{r}
invResPlot(model)
```

```{r}
summary(powerTransform(model))
```

```{r}
model2 <- lm(I(PinchingForce^(0.5)) ~ ., data = df1)

par(mfrow = c(2,2))

plot(model2)
```

transformation of $\sqrt{y}$ is more valid than untransformed model.

```{r}
pt_pred <- powerTransform(
  cbind(ThoraxLength,
        Weight,
        ClawLength,
        ClawWeight,
        ClawHeight) ~ 1,
  data = df1
)

summary(pt_pred)
```

```{r}
model3 <- lm(I(PinchingForce^(0.5)) ~ log(ThoraxLength) + log(Weight) + log(ClawLength) + log(ClawWeight) + log(ClawHeight), data = df1)

par(mfrow = c(2,2))
plot(model3)
```

```{r}
model4 <- lm(PinchingForce ~ log(ThoraxLength) + log(Weight) + log(ClawLength) + log(ClawWeight) + log(ClawHeight), data = df1)

par(mfrow = c(2,2))
plot(model4)
```

```{r}
model5 <- lm(PinchingForce ~ log(ThoraxLength) + I(log(ThoraxLength)^2) +
                           log(Weight) + I(log(Weight)^2) +
                           log(ClawLength) + I(log(ClawLength)^2) +
                           log(ClawWeight) + I(log(ClawWeight)^2) +
                           log(ClawHeight) + I(log(ClawHeight)^2),
             data = df1)

par(mfrow=c(2,2))
plot(model5)
```

```{r}
summary(powerTransform(model5))
```



```{r}
model6 <- lm(log(PinchingForce) ~ log(ThoraxLength) + I(log(ThoraxLength)^2) +
                           log(Weight) + I(log(Weight)^2) +
                           log(ClawLength) + I(log(ClawLength)^2) +
                           log(ClawWeight) + I(log(ClawWeight)^2) +
                           log(ClawHeight) + I(log(ClawHeight)^2),
             data = df1)

par(mfrow=c(2,2))
plot(model6)
```
```{r}
model7 <- lm(I(PinchingForce^(0.5)) ~ log(ThoraxLength) + I(log(ThoraxLength)^2) +
                           log(Weight) + I(log(Weight)^2) +
                           log(ClawLength) + I(log(ClawLength)^2) +
                           log(ClawWeight) + I(log(ClawWeight)^2) +
                           log(ClawHeight) + I(log(ClawHeight)^2),
             data = df1)

par(mfrow=c(2,2))
plot(model7)
```
```{r}
summary(model7)
```

```{r}
vif(model7)
```

`model3` is best right now

```{r}
boxcox(model3, lambda = seq(-0.5, 1, 0.1))
```

```{r}
model8 <- lm(I(PinchingForce^(1/3)) ~ log(ThoraxLength) + log(Weight) + 
                   log(ClawLength) + log(ClawWeight) + log(ClawHeight), 
                 data = df1)

par(mfrow=c(2,2))
plot(model8)
```

### (a)

```{r}
summary(model8)
```

Response transform: ($1/3$)

Estimated Parameters:

`Intercept`
  - -6.89845
  - no transform
`ThoraxLength`
  - 0.70178
  - log transform
`Weight`
  - -0.25851
  - log transform
`ClawLength`
  - 1.18128
  - log transform
`ClawWeight`
  - 1.57758
  - log transform
`ClawHeight`
  - -0.06667
  - log transform
  
### (b)

Trend component, both are equally as bad. Maybe mine is slightly better but not considerably. My model has a mild U-shape, while your model has a wave / S-shape that changes direction twice. Your trend suggests a more complex and unmodelled structure compared to my slight U-shape.

### (c)

My model's middle points track the line very tightly.

### (d)

Constant variance. Neither of our plots have a fanning shape in the residual plot.

Mine is better in the scale-location plot, where my line is very flat, while yours has more variation in the line.

## (2)

### subsets

```{r}
best_fit <- regsubsets(
  I(PinchingForce^(1/3)) ~ log(ThoraxLength) + log(Weight) + 
    log(ClawLength) + log(ClawWeight) + log(ClawHeight),
  data = df1,
  nvmax = 5
)
best_summary <- summary(best_fit)
best_bic_index <- which.min(best_summary$bic)
best_summary$which[best_bic_index, ]
```


### backward

```{r}
backward_fit <- regsubsets(
  I(PinchingForce^(1/3)) ~ log(ThoraxLength) + log(Weight) + 
    log(ClawLength) + log(ClawWeight) + log(ClawHeight),
  data = df1,
  method = "backward",
  nvmax = 5
)
backward_summary <- summary(backward_fit)
backward_bic_index <- which.min(backward_summary$bic)
backward_summary$which[backward_bic_index, ]
```

### forward

```{r}
forward_fit <- regsubsets(
  I(PinchingForce^(1/3)) ~ log(ThoraxLength) + log(Weight) + 
    log(ClawLength) + log(ClawWeight) + log(ClawHeight),
  data = df1,
  method = "forward",
  nvmax = 5
)
forward_summary <- summary(forward_fit)
forward_bic_index <- which.min(forward_summary$bic)
forward_summary$which[forward_bic_index, ]
```

```{r}
best_model <- lm(I(PinchingForce^(1/3) ~ log(ClawWeight)), data =df1)

summary(best_model)
```

Best model is 

$\text{PinchingForce}^{\frac{1}{3}} = \beta_0 +\beta_1log(\text{ClawWeight})$

(best subsets regression, backward stepwise, and forward stepwise regression all unanimously agreed that only including these two would lead to the best model. (used BIC))

## (3)

```{r}
best_model_sex <- lm(I(PinchingForce^(1/3) ~ log(ClawWeight) + Sex), data =df)

summary(best_model_sex)
```

```{r}
confint(best_model_sex, "SexM", level = 0.95)
```

Controlling for ClawWeight, there is no statistically significant difference in pinching force between male and female crabs (p = 0.457 >> 0.05). The estimated difference is small (males scoring 0.084 higher on the cube root scale), and the 95% confidence interval of (-0.146, 0.313) contains 0, confirming we cannot rule out no difference at all. There is insufficient evidence that Sex has any meaningful effect on PinchingForce.

## (4)

```{r}
lev <- hatvalues(best_model)

n <- nrow(df1)

threshold <- 4 / n

which(lev > threshold)

```

observations 1 and 28 in rows 1 and 22, respectively have high leverage.


```{r}
cooks <- cooks.distance(best_model)

which(cooks > threshold)
```

observations 7 and 27 in rows 7 and 21, respectively have high influence.